### 2011-12-19 joint action and knowing others minds/interacting mindreaders.2.tex
communication bit first attempt done

As this example indicates,
exploiting your-goal-is-my-goal 
can shift the burden of identifying goals from a mindreader to her target.
In the example Katarina is the focal mindreader and we are her target;
but her success in identifying the goal of our actions
depends on this,
that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
But interacting mindreaders who rely on your-goal-is-my-goal
thereby rely on their targets' having correctly identified the goals of their actions.
Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
But where the target has better understanding of relevant means-ends relations,
such as actions involving novel tools,
the your-goal-is-my-goal route to knowledge of other's goals is sometimes the only option.

@@

\section{Communicative gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we suggested that it could sometimes affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which participants had to find a reward 
and were helped either by a failed reach or by a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which occurred first and did help.
Taking this paradigm as our case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transation from 
a simple understanding of goals 
to 
understanding of communicative intent.

>>>>

As this example indicates,
exploiting your-goal-is-my-goal 
can shift the burden of identifying goals from a mindreader to her target.
In the example Katarina is the focal mindreader and we are her target;
but her success in identifying the goal of our actions
depends on this,
that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
But interacting mindreaders who rely on your-goal-is-my-goal
thereby rely on their targets' having correctly identified the goals of their actions.
Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
But where the target has better understanding of relevant means-ends relations,
such as actions involving novel tools,
the your-goal-is-my-goal route to knowledge of other's goals may sometimes be the only option.

@@

\section{Communicative gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we saw that it could  affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which chimpanzees had to find a reward 
and were helped by being shown either a failed reach or  a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
Taking this paradigm as a case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transition from 
a simple understanding of goals 
to 
an early understanding communicative actions.%
\footnote{
Please note that, although our discussion borrows an experimental paradigm
our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
that your-goal-is-my-goal
could \emph{in principle} enable
individuals to make the transition 
from
 a simple understanding of goals
to
 an early understanding of communicative actions.
Of course it would strengthen our argument if we could provide evidence to show that this actually happens.
But for now we are concerned with more narrowly conceptual issues.
}

Let us imagine ourselves as the chimpanzee  for a moment.
We witness the pointing action.
With our eyes we follow the point to a container \citep[see][p.\ 6]{Moll:2007gu}.
So we do associate the pointing action with its target.
But we are no more likely to choose this container than the other in seeking the reward.
So we probably do not think of the pointing action as having any goal which would clue us in to the relevance of the container it indicates.
(In principle we might perfectly understand the pointing action while failing to react to it in any systematic way because we are uncertain about the agent's integrity; but let us discount this possibility for the sake of illustration.)
Now suppose 
that, 
before pointing, the agent had used engaging facial gestures to signal willingness to engage in joint action with us; and
that
we were able to think of retrieving the food as a possible distributive goal%
\footnote{
    On \emph{distributive goal} see section \vref{sec:joint_action}.
  }
of our actions;
and 
that we had exploited the my-goal-is-your-goal inference.
Then we would believe,
perhaps mistakenly,
that a goal of the pointing action was to retrieve the food.
In which case the pointing action would have been no less helpful in enabling us to succeed than the failed reach---which, as you may recall, was very helpful.
So the your-goal-is-my-goal inference can 
enable interpreters
to  
misunderstand pointing actions as something like  failed reaches rather than as communicative gestures.
This means that,
even without any understanding of communication,
they can respond appropriately to helpful pointing actions in the context of joint action.

It is natural to suppose that the difficulty chimpanzees have in Hare and Tomasello's experiments with responding appropriately to helpful pointing but not to failed reaching
is due to a failure to understand communicative intention.%
\footnote{
See  footnote \vref{fn:communicative_intention}.
Relatedly,
in their discussion of these findings
Moll and Tomasello suggest that
`to understand pointing, the subject needs to understand more than the individual goal-directed behaviour. 
She needs to understand that by pointing towards a location, the other attempts to communicate to her where a desired object is located; that the other tries to inform her about something that is relevant for her'
\citep[p.\ 6]{Moll:2007gu}.
Assuming this is right, our suggestion is that individuals could reliably  respond  appropriately to pointing actions in the context of joint action without understanding pointing.
}
What we are suggesting is that participants must also have been unable or unwilling to 
exploit the your-goal-is-my-goal inference.

Consider a related experiment by \citet{leekam_adults_2010}.
Again participants had to retrieve a reward from one of several closed containers, but this time they were two- and three-year-old children.
In one condition participants were shown an adult holding up a replica of the target container.
Leekam and colleagues found that 
when this action was accompanied by an engaging facial expression,
three-year-old children were significantly better at identifying the correct container compared to
when the the action was accompanied by a neutral facial expression (p.\ 116).
Why did the engaging facial expression enhance performance?
The authors consider the idea that 
engaging facial gestures somehow help children to understand communicative intentions.%
\footnote{
\citet[p.\ 118]{leekam_adults_2010}: `the adult’s social cues conveyed her communicative intent, which in turn encouraged the child to ‘see through the sign’ ... helping them  to take a dual stance to it.'
}
An alternative possibility is that children succeeded without understand the replica as a sign at all.
Instead they may have 
associated the replica with the container it resembled
(which by itself is not enough to motivate selecting this container, of course),
regarded the engaging facial gestures as expressing willingness to engage in joint action,
and 
exploited your-goal-is-my-goal 
	to infer that a goal of the action of holding up the replica was to find the reward.
In this way they might have understood (or misunderstood) the action of holding up the replica as 
like a failed reach
in being an attempt to retrieve the reward.

So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
The basic requirement is this: in a particular context, the interpreter must associate a communicative gesture with its referent.
For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking for an object she must associate an utterance of `cupboard' with the nearby cupboard.
As we saw, 
outside the context of joint action,
merely associating a gesture with its referent falls short of being able to respond appropriately.
After all, the goals of the gesture may well be unrelated to the goals of the interpreter's actions.
But if an interpreter supposes that her target is willing to engage in joint action with her,
then she may infer that the goal of her target's action is her goal
and so be motivated to treat the thing associated with a communicative gestures as relevant to the goal of her own actions.
This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
she may be in a position to realise, further, that the same tool can be used in other contexts.

This, in barest outline, is 
how
abilities to engage in joint action
mean that 
a mindreader with an ability to ascribe simple goals only (and no understanding of communicative intent)
might 
nevertheless reliably respond appropriately to some communicative gestures,
and so be in a position to understand how such gestures can be used to guide others' actions.%
\footnote{
Contrast Csibra's claim that,
early in human development,
goal ascription (`teleological understanding' in his terms)
and 
identifying the referents of communicative gestures (`referential understanding')
`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
and
are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
%(although of course they come together later in development)
We have not shown that this hypothesis wrong.
But we have shown that there is another possibility:
the referential stance may 
emerge from
the teleological stance
 together with abilities to engage in simple forms of joint action.
}
Of course we have only argued that this transition is theoretically possible;
we have not attempted to defend any hypothesis about evolution or development.
Our point is just to illustrate one of several ways in which
interaction may matter for mindreading.
Widening the evidential basis for goal ascription
to include evidence available only to interpreters capable of joint action with their targets
makes possible 
novel hypotheses about the emergence, in evolution or development (and perhaps even both), of communicative gestures.

%This is one illustration of how capacities for joint action, even very simple forms of joint action, might be relevant to explaining the development or evolution of richer forms of cultural cognition.
